---
title: "Data Preparation"
author: "Sei Sar Hla Kyi, Sathvika Subramanian, Dew Stella Chan"
date: "2025-03-10"
date-modified: "last-modified"
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

The following document details the data preparation process for the building the prototype for the shiny app of this project assignment.

##  Libraries Used for Data Preparation. 

```{r}
pacman::p_load(tidyverse, readxl)
```

## Importing data sets

The climate data set was being scrapped from [Meteorological Service Website, Singapore](https://www.weather.gov.sg/climate-historical-daily/) using beautiful soup python package.

The scrapped historical daily records weather in Singapore were then combined using python into one csv file and loaded in R Studio.

```{r}
weather<- read_csv("data/combined_weather_data.csv")
```

```{r}
summary(weather)
```

The dengue data sets were provided at the courtesy of the National Environment Agency (NEA), Singapore.

The data set provided by NEA is also publicly available [NEA website](https://www.nea.gov.sg/dengue-zika/dengue/dengue-cases) and [Ministry of Health website.](https://www.moh.gov.sg/others/resources-and-statistics/infectious-disease-statistics-2024-weekly-infectious-diseases-bulletin)

```{r}
Weekly_dengue <-read_xlsx("data/Denguecases.xlsx", "Weekly")

Monthly_dengue <-read_xlsx("data/Denguecases.xlsx", "Monthly")
```

```{r}
summary(Weekly_dengue)
```

```{r}
summary(Monthly_dengue)
```

```{r}

weather$Mean_Temperature_C <- as.numeric(weather$`Mean Temperature (°C)`)
weather$Maximum_Temperature_C <- as.numeric(weather$`Maximum Temperature (°C)`)
weather$Minimum_Temperature_C <- as.numeric(weather$`Minimum Temperature (°C)`)


weather_Monthly <- weather %>%
  group_by(Year, Month) %>%
  summarize(
    AvgMeanTemp = mean(Mean_Temperature_C, na.rm = TRUE),
    MaxTemp = max(Maximum_Temperature_C, na.rm = TRUE),
    MinTemp = min(Minimum_Temperature_C, na.rm = TRUE)
  )
```

```{r}

Monthly_dengue_longer <- Monthly_dengue %>%
  pivot_longer(
    cols = 2:12,
    names_to = "Year",
    values_to = "denguecases"
  )%>%filter(!str_detect(`Month/Year`, "Total")) %>%
  rename(Month = `Month/Year`)


Monthly_dengue_longer$Year <- as.numeric(Monthly_dengue_longer$Year)
Monthly_dengue_longer$Month<- as.numeric(Monthly_dengue_longer$Month)
print(Monthly_dengue_longer)

```

Combining of the temperature and dengue case data.

```{r}
combined_data <- Monthly_dengue_longer %>%   left_join(weather_Monthly, by = c("Year", "Month"))
```

Exporting processed data frames to csv files to be visualised in Shiny App.

```{r}

write.csv(Monthly_dengue_longer, file = "C:/dewschan/Grp7_Shiny/data/Monthly_dengue_longer.csv", row.names = FALSE)

write.csv(weather_Monthly, file = "C:/dewschan/Grp7_Shiny/data/weather_Monthly.csv", row.names = FALSE)

write.csv(combined_data, file = "C:/dewschan/Grp7_Shiny/data/combined_data.csv", row.names = FALSE)


```
